---
title: "PopGenDataSummaryFiguresCode"
date: Dec 16 2021
output: html_document
---

Code used to generate Figure 1.

#Workspace Setup
##Load packages
```{r setup, include=FALSE}
library(readxl)
library(stringr)
library(tidyr)
library(tidyselect)
library(dplyr)
library(ggplot2)
library(webr)
library(moonBook)
library(ggforce)
library(grid)
library(RColorBrewer)
library(gridExtra)
source("PieDonutFunctions.R") #load slightly edited piedonut function https://github.com/etoff/PieDonutEMT


```


#Figure 1: Taxonomic breakdown of species represented in the CaliPopGen database. Values in parentheses represent the total number of species as a percentage of the number of unique species in the database.
Note that group labels in this plot collide, so EMT manually moved labels in post-processing using adobe illustrator. 

##Read-in database
```{r}
mydata<-read.delim("2021-12-14_PopGen_SciData.tsv", header = T, na.strings =  "<NA>")

```
##Tidy, summarize and plot data
```{r}
entrybyspecies<- mydata %>% 
  group_by(Kingdom, TaxonGroup) %>%
  summarise(n = n_distinct(EntryID))
entrybyspecies$prop<-(entrybyspecies$n/sum(entrybyspecies$n)*100)

# Get count of unique species for each Kingdom
speciesbynewtaxonomy.king<-  mydata %>% 
    group_by(Kingdom, TaxonGroup) %>%
    summarise(n=n_distinct(ScientificName))

# calculate percentage
speciesbynewtaxonomy.king$Percentage = round(100*(speciesbynewtaxonomy.king$n/sum(speciesbynewtaxonomy.king$n)), digits = 1)

# make a new TaxonGroup column and condense small groups to "Other" category
speciesbynewtaxonomy.king$TaxonGroupPlot<-speciesbynewtaxonomy.king$TaxonGroup
speciesbynewtaxonomy.king$TaxonGroupPlot[speciesbynewtaxonomy.king$Percentage < 1 ] <- "Other"

# re-summarize
speciesbynewtaxonomy.king<-speciesbynewtaxonomy.king %>%
      group_by(Kingdom, TaxonGroupPlot) %>%
      summarise(Percentage=sum(Percentage), n = sum(n))

# make nice plotting words
speciesbynewtaxonomy.king$group.perc<-paste0(speciesbynewtaxonomy.king$TaxonGroupPlot, ": ", speciesbynewtaxonomy.king$Percentage, "%" )

# sort for plotting order
speciesbynewtaxonomy.king<-arrange(speciesbynewtaxonomy.king, TaxonGroupPlot, Percentage)
#speciesbynewtaxonomy.king$group.perc<-factor(speciesbynewtaxonomy.king$group.perc, levels = speciesbynewtaxonomy.king$group.perc)


```
##Plot
```{r, echo = F}
#pdf(file = "Figure1.raw.pdf")
PieDonut2(speciesbynewtaxonomy.king, mapping = aes(pies = Kingdom, donuts =TaxonGroupPlot, count = n ), labelposition=1, showRatioThreshold = 0, donutLabelSize = 2, pieLabelSize = 3, showPieName =  F, showRatioDonut = T, labelpositionThreshold =1, ratioByGroup = F, explodeDonut = F, color = 'white', pieCols = c("#377EB8", "#999999", "#A65628", "#4DAF4A"))
#dev.off()

#Note that group labels in this plot collide, so EMT manually moved labels in post-processing using adobe illustrator. 

```
