---
title: "2022-01-11_CaliPopGen_GitHub_Figure4"
author: "JBeninde"
date: "11 01 2021"
output: html_document
---

```{r setup, include=FALSE}
library(readr)
library(rgdal)
library(raster)
library(rgeos)
library(sp)

```


## Load CaliPopGen and spatial data

Additional data is necessary to run this code and can be downloaded at URLs provided below. To run this code, download this data along with CaliPopGen dataset 1 & 2 into the same folder and replace the path in the code chunks below.

The California state border is provided by the California Government Operations Agency: https://data.ca.gov/dataset/e212e397-1277-4df3-8c22-40721b095f33/resource/3db1e426-fb51-44f5-82d5-a54d7c6e188b/download/ca-state-boundary.zip

The ecoregions for California are available here: https://databasin.org/datasets/81a3a809a2ae4c099f2e495c0b2ecc91/

Please note that the files created here were further postprocessed using QGIS and affinity photo to create Figure 4 as it appears in the publication. 

```{r}

popgen_raw =  read.delim("C:/CaliPopGen/2021-12-14_PopGen_SciData.tsv")

popgen = popgen_raw
popgen = popgen[!is.na(as.numeric(popgen$LatitudeDD)),]
popgen = popgen[!is.na(as.numeric(popgen$LongitudeDD)),]
popgen$CitationID_PopName  = paste0(popgen$CitationID, popgen$PopName)
popgen = popgen[!duplicated(popgen$CitationID_PopName),]
popgen$LatitudeDD = as.numeric(popgen$LatitudeDD)
popgen$LongitudeDD = as.numeric(popgen$LongitudeDD)
coordinates(popgen)<-~LongitudeDD+LatitudeDD
proj4string(popgen) <- crs("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")

pairwise_raw = read.delim("C:/CaliPopGen/CaliPopGen_dataset_2_population_genetic_differentiation_TSV.tsv")

CA_border = readOGR("C:/CaliPopGen/ca-state-boundary/CA_State_TIGER2016.shp")
CA_border <- spTransform(CA_border, CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"))
CA_border_UTM <- spTransform(CA_border, CRS("+proj=utm +zone=10N +datum=WGS84"))

raster_20km <- raster(xmn = bbox(CA_border_UTM)[1]-250000, xmx = bbox(CA_border_UTM)[3], ymn = bbox(CA_border_UTM)[2], ymx = bbox(CA_border_UTM)[4], res = c(20000,20000))

ecoregion_GDB = "C:/CaliPopGen/USDA Ecoregion Sections, California/data/v10/ecoregionscalifornia07_3.gdb"
GDB_list = ogrListLayers(ecoregion_GDB) 
ecoregions = readOGR(dsn=ecoregion_GDB,layer="EcoregionsCalifornia07_3_Sections_CAClip") 

```


## All populations of dataset 1 & 2 for panel A of Figure 4

```{r}

pairwise_pop1 = pairwise_raw[,c("Pop1LongitudeDD","Pop1LatitudeDD")]
pairwise_pop2 = pairwise_raw[,c("Pop2LongitudeDD","Pop2LatitudeDD")]
popgen_pop = popgen_raw[,c("LongitudeDD","LatitudeDD")]
colnames(pairwise_pop1) = c("long","lat")
colnames(pairwise_pop2) = c("long","lat")
colnames(popgen_pop) = c("long","lat")

all = rbind(pairwise_pop1, pairwise_pop2, popgen_pop)
all$long = as.numeric(all$long)
all$lat = as.numeric(all$lat)
all = all[complete.cases(all$long & all$lat),]

write.csv(all, "C:/CaliPopGen/all_populations.csv")

```


## Quantify number of populations per ecoregion for panel B of Figure 4

```{r}

ecoregions <- spTransform(ecoregions, CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"))

ecoregions_df = as.data.frame(ecoregions)
ecoregions_df$species = NA
ecoregions_df$populations = NA

for (i in 1:nrow(ecoregions_df))
{
temp_poly = ecoregions@polygons[[i]]
temp_poly = SpatialPolygons(list(temp_poly))
temp_poly = as(temp_poly, "SpatialPolygonsDataFrame")
proj4string(temp_poly) <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"
temp_data = popgen[temp_poly,]
ecoregions_df[i,4] = length(unique(temp_data$ScientificName))
ecoregions_df[i,5] = length(unique(temp_data$PopName))
}

```


## Create raster for panel C of Figure 4 (unique populations of dataset 1)

```{r}

popgen <- spTransform(popgen, CRS("+proj=utm +zone=10N +datum=WGS84"))
popgen_df = as.data.frame(popgen)

popgen_raster = rasterize(x=popgen_df[,c("LongitudeDD","LatitudeDD")], y=raster_20km,   field=popgen_df[,"CitationID_PopName"], fun='count')

writeRaster(popgen_raster,"C:/CaliPopGen/popgen_raster.asc")

```


## Create raster for panel D of Figure 4 (pairwise lines between populations of dataset 2)

```{r}

pairwise$Pop1LatitudeDD = as.numeric(pairwise$Pop1LatitudeDD)
pairwise$Pop1LongitudeDD = as.numeric(pairwise$Pop1LongitudeDD)
pairwise$Pop2LatitudeDD = as.numeric(pairwise$Pop2LatitudeDD)
pairwise$Pop2LongitudeDD = as.numeric(pairwise$Pop2LongitudeDD)

pairwise = as.data.frame(pairwise)

raster_lines = raster_20km
raster_lines[]=0

for (i in 1:nrow(pairwise))
 {
  
  line_obj <- Line(as.data.frame(matrix(c(pairwise[i,12],pairwise[i,18],pairwise[i,11],pairwise[i,17]), 2,2)))
  lines_obj <- sp::Lines(list(line_obj),ID=1)
  firstLine <- sp::SpatialLines(list(lines_obj))
  proj4string(firstLine) <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"
  firstLine_UTM <- spTransform(firstLine, CRS("+proj=utm +zone=10N +datum=WGS84"))
  raster_temp = rasterize(x=firstLine_UTM, y=raster_20km, field=1, background = 0)
  raster_lines = raster_temp + raster_lines
  
}

raster_lines[raster_lines==0] = NA
writeRaster(raster_lines,"C:/CaliPopGen/PW_raster_lines.asc")

```

